### 2013-05-10 writeup/paper.tex
adding stuff to the conclusion

\documentclass[10pt, conference, compsocconf]{IEEEtran}

%% Custom TeX

\usepackage{listings}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{amssymb}
\usepackage{url}
\usepackage[splitrule]{footmisc}

\usepackage[dvipsnames]{xcolor}

\usepackage{listings}
\lstloadlanguages{Ruby,Java}
\lstset{
basicstyle=\ttfamily\color{black},
commentstyle = \ttfamily\color{red},
keywordstyle=\ttfamily\color{blue},
stringstyle=\color{orange}}

%% for comments
\newcommand{\comment}[3][\color{red}]{{#1{[{#2}: {#3}]}}}
\newcommand{\code}[1]{\textsf{\small #1}}
\newcommand{\kris}[1]{\comment[\color{orange}]{km}{#1}}

\newcommand{\thickhline}{\noalign{\hrule height 1pt}}

%% End of custom TeX

\begin{document}
%
% --- Author Metadata here ---
% --- End of Author Metadata ---

\title{MapReduce to My Ears\\
{\Large Using Big Data for Music Synthesis}
}

%%\numberofauthors{3}
\author{
\IEEEauthorblockN{Christopher Imbriano, Amanda Strickler, Kristopher Micinski}
\IEEEauthorblockA{Computer Science Department, University of Maryland,
  College Park, MD 20742, USA
}
}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}
\label{sec:introduction}

The big data revolution has revolutionized the way we think about,
design, and implement algorithms at scale.  As such, problems from all
different areas of computer science have been lifted into the
MapReduce framework.  By contrast, MapReduce has been used relatively
less to explore art or creative tasks.  Our project looks to answer:
what kinds of art can we create harnessing large amounts of data at
scale?

We present a system to synthesize music using large amounts of data.
The input is a large corpus of images, which is subsequently processed
by our system, and the result is a musical score.  We then mix and
listen to the music.

\subsection{Motivation}

We are currently unaware of any projects using MapReduce to create art
or any other representation of digital media.  Because of this, we
anticipate that there might be problems in handling media (buffering,
representing, etc..).  We anticipate that exploring this area might
open up potential applications.

\section{Overview}
\label{sec:overview}

The core idea in converting image data to music is to view the
information in an image (color, size, opacity, etc..) as a mapping to
notes and durations with which those notes are sustained.  In our
system we chose MIDI as the output format.  Our system can then be
modeled as a function from image data to MIDI output.  Note that the
choice of MIDI limits the expressivity of our music.  Going to other
formats (such as MP3 or OOG) could provide a larger range of
primitives with which to construct music, but does not enjoy the
simplicity that MIDI provides.

Our system takes in ordered sets of images and produces a MIDI file.
In our view, a MIDI file consists of note and duration pairs.  Taken
together, these constitute a score by chaining together multiple pairs
over the course of time:

\[
MIDI : (Note,Duration)^{\ast}
\]

However, producing a MIDI score with only a single instrument would
sound broing (akin to a song with only one player): instead we take
pieces of images and map them into different parts.  Each of these
parts constitutes an instrument and --- when mixed together ---
produces a final score.  An illustration of the core idea is shown in
Figure~\ref{fig:images-to-music}.

\begin{figure*}
  \centering
  \begin{minipage}[t]{.3\textwidth}
    \centering
    \includegraphics[width=.7\textwidth]{images-to-regions.png}
  \end{minipage}
  \begin{minipage}[t]{.1\textwidth}
    \raisebox{.8 in}{\Huge $\to$}
  \end{minipage}
  \begin{minipage}[t]{.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{parts-and-instruments.png}
  \end{minipage}
  \caption{Mapping a stream of three images to a nine piece musical
    score.  First each image is mapped into regions, which will be
    transformed into notes and combined into a score using a mixer.}
\end{figure*}

\begin{figure}[h]
  \centering
  \includegraphics[width=.4\textwidth]{hsv.png}
  \caption{The HSV color cylinder}
  \label{fig:hsv}
\end{figure}

\subsection{The HSV format}

To convert image data to musical notes, we need to establish a
bijection between the data in the image and the range of music in the
song.  The way we do this is inspired by viewing colors in images as
HSV values.  In HSV, each color has three components: a hue, a
saturation, and a value.  These components are visualized (as shown in
Figure~\ref{fig:hsv}) as a cylinder (as opposed to RGB, which views
the color space as a cube).  Each component represents:

\begin{enumerate}
\item Hue: The actual ``color'' that would be described.  At full
  saturation and value, the hue determines the color seen.  This is
  visualized as going around a wheel of colors.

\item Saturation: The degree to which the color is saturated from
  white.  Zero saturation is a blank (white) surface, while increasing
  values show more permeance of color.

\item Value: The amount of light in the color.  Colors with low values
  are dark, and those with higher values are brighter.  This is
  visualized by going up the cylinder.
\end{enumerate}

In our transformation to music, we map colors in HSV to notes in MIDI.
We do this by overlaying the circle of fifths (a circle representing
pleasing intervals called fifths in music theory) onto the circle
representing hue in HSV.  We establish the octave of the produced note
using the value of the color in HSV, and the volume of the note using
the saturation.

\subsection{Generating music}
\label{sec:generating}

To generate music, we take each image in our corpus and view it as a
single step in time.  Musical scores are made by streaming large
amounts of images.  (In fact, one possible idea is to use image data
from subsequent frames in movies to get streams of related images.)
Because our music would be relatively uninteresting if it had only one
``player'', we also include several \emph{regions} in the generated
music.  Each region is a distinct musical voice that can play
different notes and is mixed together with the rest of the music.  To
state this formally, each image is mapped to a set of regions and
notes:

\[
    f : A_{i,j} \to note^n
\]

where $i$ and $j$ are the dimensions of the image, and $n$ is the
number of distinct regions of the image.  The obvious idea is to
divide the image into $n$ distinct quadrants of equal size and project
the colors in that region to notes.  One problem is that we often deal
with relatively large images: if we have a smaller set of quadrants
how do we take this to a HSV value?  We solve this problem by taking
thumbnails of images, essentially doing a projection from $A_{i,j}$
down to $A_{n^{.5},n^{.5}}$.  This also tacitly stipulates that we
start out with square images (otherwise we lose some of the image data
in the projection): we handle this step as a first preprocessing phase
where we take images and project them to square images.

In the case of MIDI, notes are pairs of integers in the range
$[-1,127]$.  The first element of the pair represents the \emph{tone}
, and the second element of the pair represents the \emph{velocity}
(how loud the note is):

\[
    note : tone \times velocity
\]

where a velocity of $-1$ represents the lack of any tone, and $127$
represents the loudest possible tone.  Colors start at $C_0$ and move
up over ten octaves.

We represent each region of the image by a representative color, which
is obtained (as noted previously) using thumbnailing.

\section{Implementation}

\begin{figure*}
  \centering
  \includegraphics[width=.85\textwidth]{pipeline.png}
  \caption{High level view of the pipeline used for the project}
  \label{fig:project-pipeline}
\end{figure*}

Our implementation consists of scripts to transform the input,
MapReduce phases, and a music mixing tool.  The high level pipeline is
shown in Figure~\ref{fig:project-pipeline}.  It consists of a sequence
of MapReduce jobs, along with sets of helper scripts

\subsection{Conversion of images to thumbnails}

Conversion to music happens by doing operations on colors and image
data.  To do the conversion to tone, velocity pairs, we transform
colors in the input data, meaning that we have to start with a
representation of $R$ colors for each image.  To do this
transformation, we considered using a MapReduce job: mapping over the
sets of images in our input corpus.  Ultimately we found that
converting images to thumbnails was better accomplished using scripts:
which allowed us to do other manipulation as well.  We anticipate that
if this became a problem we could have moved the image processing code
to Java and pass in our input corpus to the initial stage of
MapReduce.  Our current implementation uses the RMagick \cite{rmagick}
gem, which is a Ruby library interfacing to ImageMagick
\cite{imagemagick}.  ImageMagick includes facilities for taking images
to thumbnails.

\begin{figure}
\begin{lstlisting}[language=Ruby]
require 'rmagick'

class ThumbMaker
  # Convert set of images to thumbnails
  def self.convert_thumbs(dir_path, size)
      # ... 
      puts "Making thumbnail for..."
      Dir.glob("#{path}/*.jpg") do |fname|
        # ...
        fname = # ...
        img = Magick::Image.read(fname)[0]
        thumb = img.thumbnail(size, size)
        thumb.write(thumb_dest)
      end
    end
end

\end{lstlisting}
\caption{Relevant code from the module used for image to thumbnail
  conversion}
\end{figure}

\subsection{Conversion to intermediate representation}
\label{sec:conversion}

Image data comes in a variety of different formats: JPEG, GIF, PNG,
etc..  Dealing with the plethora of formats presents a problem for our
translation because there are many artifacts of the formats
(annotations, geotagging, compression, etc..) which are unnecessary
for our implementation.  To keep the core abstraction of images as
arrays of bytes (the abstraction described in
Section~\ref{sec:overview}) the first stage of our Hadoop
implementation\footnote{\code{src/main/phase1}} converts the surface
image representation to the sequence file format used in subsequent
processing.  Although it may appear inefficient to hold an
uncompressed image in memory, this is actually ideal for Hadoop:
because computing cycles (incurred through recompressing and sending
the images) are the cost, compared with network traffic.

The conversion in our prototype uses the java \code{ImageBuffer} class
to do image input and conversion.  This class holds requires a large
amount of memory to represent image objects, however this is fine
because we hold only a small number of images in memory at any one
time (using statics) when mapping over our input corpus.

Our implementation for this conversion happens in the
\code{WriteImagesToSequenceFile} class.  They key novelty in this
class is the conversion of image data to a \code{SequenceFile} based
representation.  A \code{SequenceFile} works well for our application
because we need to stream potentially large amounts of image data (raw
bytes).  This is fundamentally no different than creating a custom
Writable instance to serialize to a \code{SequenceFile} at a later
point, but given the simplicity of our representation we say no need
to do this.

\begin{figure}
\begin{lstlisting}[language=Java]
public class WriteImgsToSequence
{
  BufferedImage img = null;
  SequenceFile.Writer writer=null;
  FileSystem fs = 
      FileSystem.get(conf);
  FileStatus[] fss = 
    fs.globStatus(new Path(input 
      + "/*.jpg"));
  writer = SequenceFile
      .createWriter(fs, 
    conf,outPath,IntWritable.class,
    ArrayListOfIntsWritable.class);
  int counter = 0;
  for (FileStatus status : fss) {
    IntWritable image_counter = 
      new IntWritable();
    ArrayListOfIntsWritable 
     pixelData = 
      new ArrayListOfIntsWritable();
    img = ImageIO.read(
      new File(status.getPath()
         .toString().substring(5)));
    // Get RBG for each pixel 
    for(int j = 0; j<thumbs; j++) {
     for(int i = 0; i<thumbs; i++) {
        pixelData
          .add(img.getRGB(i,j));
      }
    }
    // Write to the sequence file
    image_counter.set(counter);
    writer.append(image_counter,
      pixelData);
    counter += 1;
  } 
}
\end{lstlisting}
\caption{Snippet of code from \code{WriteToSequenceFile} showing
  format used in sequence files throughout the MapReduce jobs.}
\end{figure}

\subsection{Translating to note, region pairs}

After taking the image data to an intermediate format, we must map the
image data to a set of note / region pairs for use in the MIDI.  To do
this, we map a conversion function across each of the different
portions of each image.  The conversion happens by mapping over our
set of images in the \code{Image2MusicMR} class.  The output of the
image to music conversion is $n$ sets of note lists (where notes are
expressed as pairs).

The strategy used in the conversion is to map over each pixel in an
image, apply the conversion function, and emit it to a reducer to
collate together.  After mapping over each pixel, we need to end up
with a stream of images for each region: a na\"{i}ve implementation simply
maps each pixel in our input corpus to these tone, velocity pairs:

\[
  map : id \times (r,g,b) \to region \times (tone \times velocity)
\]

However, when we actually mix the music, we need each of the streams
to be in an ordered format.  Because the default sorting order does
not take this into account, the music ends up jumbled (each of the
regions is permuted with respect to input ID in an arbitrary way).  To
fix this, we perform secondary sorting on the input ID:

\[
  map : id \times (r,g,b) \to (id \times region) \times (tone \times velocity)
\]

A snippet of our mapper is shown in
Figure~\ref{figure:colortomusic-mapper}.

\begin{figure}
\label{figure:colortomusic-mapper}
\begin{lstlisting}[language=Java]
public void map(IntWritable key, 
    ArrayListOfIntsWritable pixels,
    Context context)
{
  int regionCounter = 1;
  MidiNote midiNote = null;
  for(int i=0; i<pixels.size();i++){
    pixel = pixels.get(i);
    // Conversion from color to note
    midiNote = Color2Music
      .convert(new Color(pixel));
    REGION_IMAGENO.set(regionCounter,
      key.get());
    NOTE_VELOCITY.set(midiNote.getTone(),
      midiNote.getVelocity());
    context.write(REGION_IMAGENO,
       NOTE_VELOCITY);
    regionCounter++;
  }
}
\end{lstlisting}
\caption{A portion of our mapper that takes color data to MIDI notes.
  For each pixel in the original image, the conversion in
  \code{Color2Music} is applied before setting the output key (as a
  region, number pair).}
\end{figure}

\subsection{Mixing and producing playable music}

The result of our mapreduce pipeline is a set of notes in MIDI: one
for each region in the image set.  We postprocess the MapReduce output
by taking it off HDFS using a script and converting to the MIDI format
(adding on a header about the number of parts and other small details
to be a valid MIDI file). To turn these into music that can be played,
we use mixing software (such as GarageBand) to mix the regions into
full scores and generate MP3 files.

To visualize music and colors, we take the resulting music and mix it
with a video generated from the thumbnailed images in our input
dataset.  We do this using a process similar to preprocessing: we
simply blow up the thumbnailed images to a larger size, concatenate
them together into subsequent frames of a video, and mix the video
with the music generated from our MapReduce phases.

\section{Results}

To test the system, we use a variety of different sets of images.
These images were obtained from
Flickr\footnote{\url{http://www.flickr.com/}} and \code{memebase.com},
using a public API and a Python script \cite{flickerpy} to get test
data.  We converted animated GIFs into individual frames again using a
preprocessing phase.  Our MapReduce code takes as input simply a
\code{SequenceFile} (as discussed in Section~\ref{sec:conversion}), so
we simply feed in a folder of JPG images to our preprocessing phase.

As a result, we produced various pieces of music.  Our longest piece
(named \code{Digital Nightmare}) was over six minutes long and
composed of over 3000 images.  We also produced shorter (self
contained) pieces from various related images: such as GIFs of cat
jumping animations.  All of these pieces are made available in our
repository.

Our MapReduce phases ran very quickly (on the order of minutes, mostly
startup time).  We believe that the reason for this is that by the
time preprocessing was completed we ended up with a relatively small
amount of data to be fed into MapReduce.  Were we to do the project
over again we would have fed in much more data to the MapReduce jobs
and optimized the selection of regions (from each image) to be the
``best sounding'' piece of music that could be produced.  We thought
about how to devise a measure of ``realistic'' music, but could not
easily see one.  One potential next application would be to use Hadoop
to mine this correlation.  The idea is that realistic music generally
has some recurring nature (backbeat, for example) even if very
complex.  Therefore, we would want to apply a penalty term to our
measure whenever we generated music that had too complex a principal
structure (this would avoid creating chaotic sounding music).  We did
some initial work on this (writing down some equations, studying how
to analyze MIDI for it's beat) but didn't have anything concrete by
the time we needed to do implementation.

We also noticed that there might be some interesting problems in
taking a song and a set of videos and finding out which best matched
each other.  (In essence, this would let you search for image based on
video.)  We started to write an inverse transformation within
MapReduce\footnote{src/main/phase2/Color2Music} to explore this idea,
but haven't finished doing the work yet.

\section{Related work}

Many domain specific programming languages exist for music synthesis
[\cite{haskore},\cite{lilypad}].  Many of these languages synthesize
music in the small ().  By contrast, our approach focuses on using
large amounts of data to create music...

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{paper}

\end{document}